Typed Functional Genetic Programming

### 2013-06-24 tex/thesis.tex
[tex]

@@

%For brevity we will als

@@

Satisfying this constrain ensures that every program tree build 
from $T \cup F$ will be total.

@@

\subsection{Generating initial population}
We will describe tree generating method used by Koza in \cite{koza92} called
\textit{ramped half-and-half}.

\subsection{GP algorithm in pseudocode}

@@

\Pseudokod{GP(fitnessFun,buildingBlocks,numGens,popSize,probabs)}{
	$gen \leftarrow$ 0 \;
	$pop \leftarrow generateInitialPopuletion( buildingBlocks ) $ \;
	($popWithFitnesses,terminate) \leftarrow$ evaluate($fitnessFun$, $pop$) \;

@@

	$newPop \leftarrow$ empty population \;
	$i \leftarrow 0$ \;

	\While{ $i < popSize$ }{
		$op \leftarrow probabilisticallySelectOperation(probabs)$ \;		
		\Switch{op}{
			\Case{Crossover}{
				$indiv1 \leftarrow$ selection( $popWithFitnesses$ ) \;
				$indiv2 \leftarrow$ selection( $popWithFitnesses$ ) \;
				($child1$,$child2$) = crossover( $indiv1$ , $indiv2$ ) \;
				$newPop$.insert( $child1$ ) \;
				$newPop$.insert( $child2$ ) \;
				$i \leftarrow i + 2$ \;
			}
			\Case{Reproduction}{
				$indiv \leftarrow$ selection( $popWithFitnesses$ ) \;
				$newPop$.insert( $indiv$ ) \;
				$i \leftarrow i + 1$ \;
			}
			\Case{Mutation}{
				$indiv \leftarrow$ selection( $popWithFitnesses$ ) \;
				$mutant \leftarrow$ mutate( $indiv$ ) \;
				$newPop$.insert( $mutant$ ) \;
				$i \leftarrow i + 1$ \;			
			}		
		}
	}

	$pop \leftarrow newPop$  \;
	$gen \leftarrow gen + 1$ \;
	\;
	($popWithFitnesses,terminate) \leftarrow$ evaluate($fitnessFun$, $pop$) \; 
 }
 \Return pop \;
}

@@

	\subsection{Input}
	\subsection{Output}
	\subsection{GP algorithm}
		\subsubsection{Generating initial population}
		\subsubsection{Selection}
		\subsubsection{Crossover}
		\subsubsection{Reproduction}
		\subsubsection{Mutation}

    ...
	\\\\
	\textbf{TODO}
	\begin{itemize} 
		\item TALK ABOUT GP is part of EA etc. and maybe define the GP by defining EA and then
		      specifying the differences or something like that...    
		\item History, citations, etc ....
		\item ten algoritmus v kozovi dělá "Designate Result" já tam vracim poslední populaci
	\end{itemize}

>>>>

\newcommand{\EA}{EA } % TODO nahradit správnym "evoluťionar algorithms"

\newcommand{\setDots}[2]{ 
	\lbrace #1 \dots #2 \rbrace
}

@@

Beside symbols in $T \cup F$ there must also be 
an implementation for each symbol which is not variable. 
And for every function symbol from $F$ there must be specified 
number of arguments.\\

@@

Satisfying this constrain ensures that every program tree build 
from $T \cup F$ will be total.\\

@@

\subsection{Generating initial population}

We will describe tree generating method used by Koza in \cite{koza92} called
\textit{ramped half-and-half}.

\textbf{TODO!}

\subsection{Selection}

In \EA there is plenty of options for selection mechanisms 
for us to choose from. Again we will describe mechanism Koza
used in his first book on GP. It is \textit{Roulette selection}.
It uses fitness value of each individual in straightforward way to determine probability of selecting this individual.\\

Let there be $popSize$ individuals in the population.\\
And let $f_{i}$ be fitness value for individual i 
where $i \in \setDots{1}{popSize}$.

Then $p_{i}$ probability of selection of individual $i$ is computed
as follows:\\

$p_{i} = \dfrac{ f_{i}  }{ \Sigma_{j=1}^{popSize} f_{j} } $

\subsection{Crossover}

For John Koza the most important genetic operator in GP is 
crossover. It is operator inspired by sexual reproduction
occurring in the nature. Generally speaking crossover takes
two (parent) individuals and combines theirs genomes to produce 
two possibly new (child) individuals.

In GP the most common mutation is \textit{Subtree swapping mutation}.
This crossover randomly selects one node in each parent tree.
Two new child individuals are constructed by swapping subtrees 
which have roots in those selected nodes.\\

Example should clarify this process. Here are two parent trees with 
selected nodes in bold:

\Tree [.$ifneq$ $1$
		 	   [.\textbf{iflt} $0$ $x$ [.$-$ $0$ $x$ ] $1$ ]
		 	   [.$+$ \text{$x$} \text{$2$} ]
		 	   $1$ ]
\Tree [.$\%$ \text{$x$}
         	 [.\textbf{ifeq} \text{$1$} \text{$x$} \text{$x$} \text{$0$} ] ]\\

And here are two child trees with swapped subtrees:

\Tree [.$ifneq$ $1$
		 	   [.\textbf{ifeq} \textbf{1} \textbf{x} \textbf{x} 
		 	     \textbf{0} ]
		 	   [.$+$ \text{$x$} \text{$2$} ]
		 	   $1$ ]
\Tree [.$\%$ \text{$x$}
         	 [.\textbf{iflt} \textbf{0} \textbf{x} 
         	   [.\textbf{-} \textbf{0} \textbf{x} ] \textbf{1} ] ]\\

\subsection{Reproduction}
\subsection{Mutation}

\newpage
\subsection{GP algorithm in pseudocode}

@@

\Pseudokod{GP(fitness,buildingBlocks,numGens,popSize,probabs)}{
	$gen \leftarrow$ 0 \;
	$pop \leftarrow generateInitialPopuletion( buildingBlocks ) $ \;
	($popWithF,terminate,best) \leftarrow$ evaluate($fitness$, $pop$)\;

@@

	$newPop \leftarrow$ empty population \;
	$newPop$.insert( $best$ )\;
	$i \leftarrow 1$ \;

	\While{ $i < popSize$ }{
		$op \leftarrow probabilisticallySelectOperation(probabs)$ \;		
		\Switch{op}{
			\Case{Crossover}{
				$parent1 \leftarrow$ selection( $popWithF$ ) \;
				$parent2 \leftarrow$ selection( $popWithF$ ) \;
				($child1$,$child2$) = crossover( $parent1$ , $parent2$ ) \;
				$newPop$.insert( $child1$ ) \;
				$newPop$.insert( $child2$ ) \;
				$i \leftarrow i + 2$ \;
			}
			\Case{Reproduction}{
				$indiv \leftarrow$ selection( $popWithF$ ) \;
				$newPop$.insert( $indiv$ ) \;
				$i \leftarrow i + 1$ \;
			}
			\Case{Mutation}{
				$indiv \leftarrow$ selection( $popWithF$ ) \;
				$mutant \leftarrow$ mutate( $indiv$ ) \;
				$newPop$.insert( $mutant$ ) \;
				$i \leftarrow i + 1$ \;			
			}		
		}
	}

	$pop \leftarrow newPop$  \;
	($popWithF,terminate,best) \leftarrow$ evaluate($fitness$, $pop$)\;
	$gen \leftarrow gen + 1$ \; 
 }
 \Return pop \;
}

@@

	\subsection{Input}
	\subsection{Output}

    ...
	\\\\
\subsection{TODO}
	\begin{itemize} 
		\item TALK ABOUT GP is part of EA etc. and maybe define the GP by defining EA and then
		      specifying the differences or something like that...    
		\item History, citations, etc ....
		\item ten algoritmus v kozovi dělá "Designate Result" já tam vracim poslední populaci
	\end{itemize}
